---
title: "Table 1"
output: html_document
date: "2024-07-05"
---

## Table 1 code

```{r}
df_crossectional <- read.csv("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Data/All_metrics_crossectional_samples.csv", header=TRUE, stringsAsFactors=FALSE)

df_crossectional <- df_crossectional[df_crossectional$Array == "EPICv1", ]

```

## See collumn names

You can also embed plots, for example:
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}

#Assume 'df' is your dataframe
column_names <- names(df_crossectional)

#Print the column names
print(column_names)
```


Frequency table
```{r}
# frequency_table <- table(df_crossectional$Main.Diet)
# print(frequency_table)


```

Return number of NA values

```{r}

# Calculate the number of NA values in the column
na_count <- sum(is.na(df_crossectional$Main.Diet))

# Print the number of NA values
print(na_count)

```
Return non NA values

```{r}
# Calculate the number of non-NA values in the column
non_na_count <- sum(!is.na(df_crossectional$Main.Diet))

# Print the number of non-NA values
print(non_na_count)
```

Function to see NA and Non NA values for a list of collumns

```{r}
count_na_non_na_empty <- function(df, columns) {
  # Initialize an empty list to store results
  results <- list()
  
  # Loop through each column name in the columns list
  for (column in columns) {
    # Check if the column exists in the data frame
    if (column %in% names(df)) {
      # Calculate the number of NA, non-NA values, and empty cells
      na_count <- sum(is.na(df[[column]]))
      non_na_count <- sum(!is.na(df[[column]]))
      empty_count <- sum(df[[column]] == "" & !is.na(df[[column]]))
      
      # Adjust non_na_count to exclude empty cells
      non_na_count <- non_na_count - empty_count
      
      # Store the results in the list
      results[[column]] <- data.frame(
        Column = column,
        NA_Values = na_count,
        Non_NA_Values = non_na_count,
        Empty_Values = empty_count
      )
    } else {
      # Handle the case where the column does not exist in the data frame
      warning(paste("Column", column, "does not exist in the data frame"))
    }
  }
  
  # Combine the individual data frames into a single data frame
  result_table <- do.call(rbind, results)
  
  return(result_table)
}

# List of columns to check
columns_to_check <- c('Main.Diet', 'Primary.Diet.Past.Year', 'Strictly.Followed.Main.Diet', 'Main.Diet.Factor', "Red.Meat.times.per.week","Processed.Food.times.per.week", 'Lab.Fasting', 'Nutritional.Supplements', "Rapamycin", "Metformin", "Anti.Aging.Interventions", "Supplements.Medication", "GrimAge.PC", "PhenoAge.PC", "OMICmAge", "DNAmFitAge", "FitAgeAccel", "DNAmVO2max", "SystemsAge.Blood"
                      ,"SystemsAge", "Ethnicity.Ancestry","Ethnicity", "Biological.Sex", "Alcohol.per.week", "Level.of.Education","Stress.Level","Feel.Well.Rested.days.per.week", "Tobacco.Use", "Pack.Years.Smoker", 
                      "Recreational.Drug.Use", "Recreational.Drug.Frequency","Exercise.per.week", "Exercise.Type", "Hours.of.sleep.per.night", "Metformin"                                                                                          
,"TA.65"                                                                                              
,"NR.NAD"                                                                                             
,"Sulphoraphane.broccoli.extract"                                                                     
,"Rapamycin"                                                                                          
,"DHEA"                                                                                               
,"Resveratrol"                                                                                        
,"Antiaging.Exosome"                                                                                  
,"Antiaging.Hormone.Replacement"                                                                      
,"Antiaging.Senolytics"                                                                               
,"Antiaging.Stem.Cell"                                                                                
,"Antiaging.NAD"                                                                                      
,"Recreational.Drug.Use.any"                                                                          
,"Amphetamines"                                                                                       
,"Benzodiazepines"                                                                                    
,"Marijuana"                                                                                          
,"Hallucinogens"                                                                                      
,"MDMA"                                                                                               
,"Opioids"
,"BMI"                                                                                                
,"Main.Diet.Factor"
, "Hours.Sedentary.Remaining.Awake")

# Call the function and print the result
result <- count_na_non_na_empty(df_crossectional, columns_to_check)
print(result)


```

See what type of values we have in the first 3 Main Diet, primary diet pas uear and strictly followed main diet.

### Main.Diet
```{r}

frequency_table <- table(df_crossectional$Biological.Sex)
print(frequency_table)

```

### Primary.Diet.Past.Year	
```{r}

frequency_table <- table(df_crossectional$Primary.Diet.Past.Year	)
print(frequency_table)

```

### Strictly.Followed.Main.Diet	
```{r}

frequency_table <- table(df_crossectional$Strictly.Followed.Main.Diet	)
print(frequency_table)

```

### Lab.Fasting	
```{r}

frequency_table <- table(df_crossectional$Prediabetes	)
print(frequency_table)

```

### Lab.Fasting	histograms - Glucose

You can use the ggplot2 package in R to create histograms for the glucose values grouped by the fasting status and color them accordingly. Here’s how you can do it:

First, ensure you have the ggplot2 package installed. If not, you can install it using install.packages("ggplot2").
```{r}

# Load the ggplot2 package
library(ggplot2)

# Remove NA values from the relevant columns
df_crossectional <- df_crossectional[!is.na(df_crossectional$Clinical_EBP_Glucose) & !is.na(df_crossectional$Lab.Fasting), ]

# Create the histogram
ggplot(df_crossectional, aes(x = Clinical_EBP_Glucose, fill = Lab.Fasting)) +
  geom_histogram(binwidth = 1, position = "dodge", alpha = 0.7) +
  labs(title = "Histogram of Glucose Values by Fasting Status",
       x = "Glucose Values",
       y = "Frequency",
       fill = "Fasting Status") +
  theme_minimal() +
  scale_fill_manual(values = c("Yes" = "blue", "No" = "red"))



```


### Lab.Fasting	histograms - HBA1c


```{r}
# Load the ggplot2 package
library(ggplot2)

# Remove NA values from the relevant columns
df_crossectional <- df_crossectional[!is.na(df_crossectional$Clinical_EBP_HbA1c) & !is.na(df_crossectional$Lab.Fasting), ]

# Create the histogram
ggplot(df_crossectional, aes(x = Clinical_EBP_HbA1c, fill = Lab.Fasting)) +
  geom_histogram(binwidth = 0.05, position = "dodge", alpha = 0.7) +
  labs(title = "Histogram of Clinical_EBP_HbA1c by Fasting Status",
       x = "Clinical_EBP_HbA1c",
       y = "Frequency",
       fill = "Fasting Status") +
  theme_minimal() +
  scale_fill_manual(values = c("Yes" = "blue", "No" = "red"))


```



### Lab.Fasting	histograms - triglycerides


```{r}

# Load the ggplot2 package
library(ggplot2)

# Remove NA values from the relevant columns
df_crossectional <- df_crossectional[!is.na(df_crossectional$GrimAge.PC) & !is.na(df_crossectional$Lab.Fasting), ]

# Create the histogram
ggplot(df_crossectional, aes(x = DunedinPoAm, fill = Lab.Fasting)) +
  geom_histogram(binwidth = 1, position = "dodge", alpha = 0.7) +
  labs(title = "Histogram of Glucose Values by Fasting Status",
       x = "Glucose Values",
       y = "Frequency",
       fill = "Fasting Status") +
  theme_minimal() +
  scale_fill_manual(values = c("Yes" = "blue", "No" = "red"))

```

