l---
title: "Handle outliers"
output: html_document
date: "2024-10-28"
---

### Import data and functions
```{r}
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/remove_other().R")
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/logical_to_numeric().R")
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/convert_to_factors().R")
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/convert_yes_no_to_logical().R")
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/convert_to_binary().R")
# source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/char_cols_to_factors().R")
# 

source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/count_outliers().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/get_Numeric_and_int_cols().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/plot_with_std_outliers().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/replace_zero_negatives().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/count_zero_and_negatives().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/remove_na_rows().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/remove_outliers_sd().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/create_histogram().R")
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/handle_hours_sedentary_remaining_awake().R")
load("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step2")

# Rename dataframe in this new script
data<-data_fixed_datatypes 
rm(data_fixed_datatypes) # Remove old dataframe

```

### Count zeros and negatives before removal
```{r}
amount_zero_and_negatives_before_removal<-count_zero_and_negatives(data, exclude_cols=c("Patient.ID", "PID", "survey_version", "Array", 
                                                      "Collection.Date", "Decimal.Chronological.Age", 
                                                      "Stress.Level",  
                                                      "Age.mother.at.birth", "Age.father.at.birth", 
                                                      "Hours.Sedentary.Remaining.Awake", "BMI", 
                                                      "Hannum.PCAgeDev", "Horvath.PCAgeDev", 
                                                      "Telomere.Values.PCAgeDev", "GrimAge.PCAgeDev", 
                                                      "PhenoAge.PCAgeDev", "SystemsAge.BloodAgeDev", 
                                                      "SystemsAge.BrainAgeDev", "SystemsAge.InflammationAgeDev", 
                                                      "SystemsAge.HeartAgeDev", "SystemsAge.HormoneAgeDev", 
                                                      "SystemsAge.ImmuneAgeDev", "SystemsAge.KidneyAgeDev", 
                                                      "SystemsAge.LiverAgeDev", "SystemsAge.MetabolicAgeDev", 
                                                      "SystemsAge.LungAgeDev", "SystemsAge.MusculoSkeletalAgeDev", 
                                                      "SystemsAgeAgeDev", "OMICmAgeAgeDev"))
```


### Remove rows with zero, negative or NA values in outcome columns
```{r}
# Remove rows with zero, negative or NA values in outcome columns
data_with_outliers_removed <- replace_zero_negatives(data, cols=c("Hannum.PC", "Horvath.PC",
  "Telomere.Values.PC", "GrimAge.PC", "PhenoAge.PC", "SystemsAge.Blood",
  "SystemsAge.Brain", "SystemsAge.Inflammation", "SystemsAge.Heart",
  "SystemsAge.Hormone", "SystemsAge.Immune", "SystemsAge.Kidney",
  "SystemsAge.Liver", "SystemsAge.Metabolic", "SystemsAge.Lung",
  "SystemsAge.MusculoSkeletal", "SystemsAge", "OMICmAge", "DunedinPACE"))

```


### Test removal of zeroe and negative epigenetic ages
```{r}

amount_zero_and_negatives_after_removal<-count_zero_and_negatives(data_with_outliers_removed, exclude_cols=c("Patient.ID", "PID", "survey_version", "Array", 
                                                      "Collection.Date", "Decimal.Chronological.Age", 
                                                      "Stress.Level",  
                                                      "Age.mother.at.birth", "Age.father.at.birth", 
                                                      "Hours.Sedentary.Remaining.Awake", "BMI", 
                                                      "Hannum.PCAgeDev", "Horvath.PCAgeDev", 
                                                      "Telomere.Values.PCAgeDev", "GrimAge.PCAgeDev", 
                                                      "PhenoAge.PCAgeDev", "SystemsAge.BloodAgeDev", 
                                                      "SystemsAge.BrainAgeDev", "SystemsAge.InflammationAgeDev", 
                                                      "SystemsAge.HeartAgeDev", "SystemsAge.HormoneAgeDev", 
                                                      "SystemsAge.ImmuneAgeDev", "SystemsAge.KidneyAgeDev", 
                                                      "SystemsAge.LiverAgeDev", "SystemsAge.MetabolicAgeDev", 
                                                      "SystemsAge.LungAgeDev", "SystemsAge.MusculoSkeletalAgeDev", 
                                                      "SystemsAgeAgeDev", "OMICmAgeAgeDev"))

```

### Add sedentary level based on fixed thresholds (objective)
```{r}


# Using dplyr and case_when
library(pacman)
p_load(dplyr)

data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(sedentary_level_objective = case_when(
    Hours.Sedentary.Remaining.Awake >= 1  & Hours.Sedentary.Remaining.Awake <= 6 ~ 3,
    Hours.Sedentary.Remaining.Awake >= 7  & Hours.Sedentary.Remaining.Awake <= 12 ~ 2,
    Hours.Sedentary.Remaining.Awake >= 13 & Hours.Sedentary.Remaining.Awake <= 18 ~ 1,
    Hours.Sedentary.Remaining.Awake >= 19 & Hours.Sedentary.Remaining.Awake <= 24 ~ 0,
    TRUE ~ NA_real_
  ))

```


### Add sedentary level based on quartiles
```{r}

# Function creates a new column with sedentary level based on quartiles as 2 new collumns: sedentary_level_string and sedentary_level_quartiles_numeric
data_with_outliers_removed <- handle_hours_sedentary_remaining_awake(
  data = data_with_outliers_removed,
  column_name = "Hours.Sedentary.Remaining.Awake"
)

data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(sedentary_level_quartiles_numeric)



```



### test linear relationships of hours sedentary
```{r}
# Plot the scatter plot
# plot(data_with_outliers_removed$sedentary_level, data_with_outliers_removed$DunedinPACE, main = "Scatter Plot with Regression Line",
#      xlab = "X-axis", ylab = "Y-axis",
#      pch = 19, col = "blue")

# Add regression line
# abline(lm(data_with_outliers_removed$DunedinPACE ~ data_with_outliers_removed$sedentary_level), col = "red")

# lm(data_with_outliers_removed$DunedinPACE ~ data_with_outliers_removed$sedentary_level)

```


### Codify current smoking level numerically
```{r}

# # Convert the ordered factor to numeric, starting at zero, preserving NAs
# # Check missing data for Tobacco.Use by survey version
# cat("Missing data for Tobacco.Use by survey version:\n\n")

# for (survey in unique(data_with_outliers_removed$survey_version)) {
#   survey_data <- data_with_outliers_removed %>% 
#     filter(survey_version == survey)
  
#   cat("Survey Version", survey, "(n =", nrow(survey_data), "):\n")
  
#   # Calculate missing percentage
#   missing_count <- sum(is.na(survey_data$Tobacco.Use))
#   missing_percentage <- (missing_count / nrow(survey_data)) * 100
  
#   cat("Missing values:", missing_count, "(", round(missing_percentage, 2), "%)\n")
  
#   # Show frequency table including NAs
#   cat("Frequency table:\n")
#   print(table(survey_data$Tobacco.Use, useNA = "ifany"))
#   cat("\n")
# }

# Convert the ordered factor to numeric, starting at zero, preserving NAs
data_with_outliers_removed$Tobacco.Use.Numeric <- ifelse(
  is.na(data_with_outliers_removed$Tobacco.Use),
  NA_real_,
  as.numeric(data_with_outliers_removed$Tobacco.Use) - 1
)

# Create reduced tobacco use variable with 3 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Tobacco.Use_reduced = case_when(
    Tobacco.Use == "None" ~ "Non-smokers",
    Tobacco.Use %in% c("Less than 1 cigarette per week", "Less than 1 cigarette per day", "1-5 cigarettes per day", "6-10 cigarettes per day", "11-20 cigarettes per day", "More than 20 cigarettes per day") ~ "Smokers",
    is.na(Tobacco.Use) ~ "Missing",
    TRUE ~ "Missing"
  ))

# Move the Tobacco.Use.Numeric column to the end
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Tobacco.Use.Numeric, Tobacco.Use_reduced, .after = last_col())



```

### Remove all BMI under 15
```{r}
p_load(dplyr)

# Create a new data frame with only the removed rows (BMI < 15 and not NA)
removed_low_BMI_rows <- data_with_outliers_removed %>%
  filter(BMI < 15 & !is.na(BMI))

# Create a new data frame with the remaining rows (BMI >= 15 or BMI is NA)
data_with_outliers_removed <- data_with_outliers_removed %>%
  filter(BMI >= 15 | is.na(BMI))


# Move the Tobacco.Use.Numeric column to the end
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(BMI, .after = last_col())
  
# Sort the data frame according to the BMI column for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(BMI)

# Check which rows where removed for testing
removed_low_BMI_rows <- removed_low_BMI_rows %>%
  relocate(BMI, .after = last_col())

```


### Codify education level
```{r}
# Education Level Transformation
# 
# This code segment transforms the 'Level.of.Education' column in two ways:
#
# 1. Creates a numeric representation (Education_levels_numeric):
#    - 0: Did not complete high school
#    - 1: High school or equivalent
#    - 2: Some post-secondary education (Technical certificate, Associate degree, Some college)
#    - 3: Bachelor's degree
#    - 4: Advanced degrees (Master's, PhD, Professional degrees)
#
# 2. Creates a binary categorization (education_harmonized):
#    - "bachelor_or_higher": Bachelor's degree or higher
#    - "below_bachelor": Less than a Bachelor's degree
#    - "missing": NA values or any unclassified responses
#
# The code then reorganizes the dataframe to place the education columns at the end
# and sorts the data by the numeric education level for easier verification.

# Define the mapping for the education levels
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Education_levels_numeric = case_when(
    Level.of.Education == "Did not complete high school" ~ 0,
    Level.of.Education == "High school or equivalent" ~ 1,
    Level.of.Education %in% c("Technical or occupational certificate", "Associate degree", "Some college coursework completed") ~ 2,
    Level.of.Education == "Bachelor’s degree" ~ 3,
    Level.of.Education %in% c("Master’s degree", "Doctorate (PhD)", "Professional (MD, DO, DDS, JD)") ~ 4,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))

# Create harmonized education variable
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(education_harmonized = case_when(
    Level.of.Education %in% c("Bachelor’s degree", "Master’s degree", "Doctorate (PhD)", "Professional (MD, DO, DDS, JD)") ~ "bachelor_or_higher",
    Level.of.Education %in% c("Did not complete high school", "High school or equivalent", "Technical or occupational certificate", "Associate degree", "Some college coursework completed") ~ "below_bachelor",
    is.na(Level.of.Education) ~ "missing",
    TRUE ~ "missing"  # Handles any other cases
  ))

# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Level.of.Education, education_harmonized, .after = last_col())

# Sort for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Education_levels_numeric)
```


### Codify alcohol level
```{r}
# Alcohol Consumption Data Processing and Coding
# 
# This section processes alcohol consumption data by creating numeric and categorical encodings:
# 
# 1. **Numeric Encoding**: Converts categorical alcohol consumption levels to numeric values:
#    - "Never" → 0
#    - "On special occasions" → 1  
#    - "Once per week" → 2
#    - "3-5 times per week" → 3
#    - "Regularly" → 4
#    - Missing/other values → NA
# 
# 2. **Binary Categorization**: Creates a simplified binary variable:
#    - "No alcohol": Never consume alcohol
#    - "Some alcohol": Any level of alcohol consumption
#    - "Missing": NA or unrecognized values
# 
# 3. **Data Organization**: 
#    - Relocates alcohol-related variables to the end of the dataset for easier inspection
#    - Sorts the dataset by the numeric alcohol variable for validation and testing
# 
# The code uses dplyr functions (mutate, case_when, relocate, arrange) to transform
# the original categorical alcohol consumption data into more analysis-friendly formats.
unique(data_with_outliers_removed$Alcohol.per.week)

# Load the dplyr package
p_load(dplyr)

# Define the mapping for the alcohol consumption levels
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Alcohol_per_week_numeric = case_when(
    Alcohol.per.week == "Never" ~ 0,
    Alcohol.per.week == "On special occasions" ~ 1,
    Alcohol.per.week == "Once per week" ~ 2,
    Alcohol.per.week == "3-5 times per week" ~ 3,
    Alcohol.per.week == "Regularly" ~ 4,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))

# Create reduced alcohol variable with 2 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(alcohol_reduced = case_when(
    Alcohol.per.week == "Never" ~ "No alcohol",
    Alcohol.per.week %in% c("On special occasions", "Once per week", "3-5 times per week", "Regularly") ~ "Some alcohol",
    is.na(Alcohol.per.week) ~ "Missing",
    TRUE ~ "Missing"
  ))

# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Alcohol.per.week, alcohol_reduced, .after = last_col())

# Sort for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Alcohol_per_week_numeric)

```


### Codify exercise per week
```{r}
unique(data_with_outliers_removed$Exercise.per.week)

# Load the dplyr package
p_load(dplyr)

# Define the mapping for the Exercise.per.week levels
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Exercise.per.week_numeric = case_when(
    Exercise.per.week == "Never" ~ 0,
    Exercise.per.week == "1-2 times per week" ~ 1,
    Exercise.per.week == "3-4 times per week" ~ 2,
    Exercise.per.week == "5-7 times per week" ~ 3,
    Exercise.per.week == "8 or more times per week" ~ 4,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))

# Create reduced exercise variable with 2 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(exercise_reduced = case_when(
    Exercise.per.week == "Never" ~ "No exercise",
    Exercise.per.week %in% c("1-2 times per week", "3-4 times per week", "5-7 times per week", "8 or more times per week") ~ "Some exercise",
    is.na(Exercise.per.week) ~ "Missing",
    TRUE ~ "Missing"
  ))

# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Exercise.per.week, exercise_reduced, .after = last_col())

# Sort for testing
  data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Exercise.per.week_numeric)

```



### Codify Marital.Status
```{r}
unique(data_with_outliers_removed$Marital.Status)

# Load the dplyr package
p_load(dplyr)

data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Marital.Status_numeric = case_when(
    Marital.Status == "Single" ~ 0,
    Marital.Status == "Divorced" ~ 0,
    Marital.Status == "Widowed" ~ 0,
    Marital.Status == "Separated" ~ 0,
    Marital.Status == "Partner" ~ 1,
    Marital.Status == "Married" ~ 2,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))
  
# Create reduced marital status variable with 2 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(marital_status_reduced = case_when(
    Marital.Status %in% c("Married", "Partner") ~ "In a relationship",
    Marital.Status %in% c("Single", "Divorced", "Widowed", "Separated") ~ "Not in a relationship",
    is.na(Marital.Status) ~ "Missing",
    TRUE ~ "Missing"
  ))
  
# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Marital.Status, marital_status_reduced, .after = last_col())

# Sort for testing
  data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Marital.Status_numeric)
```

### Codify Hours.of.sleep.per.night
```{r}
unique(data_with_outliers_removed$Hours.of.sleep.per.night)

# Load the dplyr package
p_load(dplyr)

data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Hours.of.sleep.per.night_numeric = case_when(
    Hours.of.sleep.per.night == "6 hours or less" ~ 0,
    Hours.of.sleep.per.night == "6 to 8 hours" ~ 1,
    Hours.of.sleep.per.night == "More than 8 hours" ~ 2,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))

# Create reduced sleep hours variable with 2 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(sleep_hours_reduced = case_when(
    Hours.of.sleep.per.night %in% c("6 hours or less", "6 to 8 hours") ~ "Less than 8 hours",
    Hours.of.sleep.per.night == "More than 8 hours" ~ "More than 8 hours",
    is.na(Hours.of.sleep.per.night) ~ "Missing",
    TRUE ~ "Missing"
  ))
  
# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Hours.of.sleep.per.night, sleep_hours_reduced, .after = last_col())

# Sort for testing
  data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Hours.of.sleep.per.night_numeric)


```

### Codify Feel.Well.Rested.days.per.week
```{r}
unique(data_with_outliers_removed$Feel.Well.Rested.days.per.week)

# Load the dplyr package
p_load(dplyr)

data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Feel.Well.Rested.days.per.week_numeric = case_when(
    Feel.Well.Rested.days.per.week == "0-1 days" ~ 0,
    Feel.Well.Rested.days.per.week == "1-3 days" ~ 1,
    Feel.Well.Rested.days.per.week == "3-5 days" ~ 2,
    Feel.Well.Rested.days.per.week == "6-7 days" ~ 3,
    TRUE ~ NA_real_  # Handles any other cases or NA values
  ))

# Create reduced well-rested variable with 2 categories
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(well_rested_reduced = case_when(
    Feel.Well.Rested.days.per.week %in% c("0-1 days", "1-3 days") ~ "3 days or less",
    Feel.Well.Rested.days.per.week %in% c("3-5 days", "6-7 days") ~ "More than 3 days",
    is.na(Feel.Well.Rested.days.per.week) ~ "Missing",
    TRUE ~ "Missing"
  ))

# Print percentage of each category
cat("=== Feel.Well.Rested.days.per.week Distribution ===\n")
freq_table <- table(data_with_outliers_removed$Feel.Well.Rested.days.per.week, useNA = "ifany")
percentages <- round(prop.table(freq_table) * 100, 2)

for (i in 1:length(percentages)) {
  category <- names(percentages)[i]
  if (is.na(category)) category <- "Missing"
  cat("  ", category, ": ", percentages[i], "% (n = ", freq_table[i], ")\n", sep = "")
}
  
# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(well_rested_reduced, .after = last_col())

# Sort for testing
  data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(well_rested_reduced)

```


### Codify Ethnicity
```{r}
unique(data_with_outliers_removed$Ethnicity)

# Load the dplyr package
p_load(dplyr)

# Define the mapping for the Caffeine.Use consumption levels
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(Ethnicity_reduced = case_when(
    Ethnicity == "Euro" ~ "European",
    Ethnicity == "Asia" ~ "Asian",
    Ethnicity == "Lati" ~ "Other",
    Ethnicity == "Afr_Ame" ~ "Other",
    Ethnicity == "Mid_afr" ~ "Other",
    Ethnicity == "Nat_ame" ~ "Other",
    Ethnicity == "Ocea" ~ "Other",
    Ethnicity == "Other" ~ "Other",
    Ethnicity == "Saha" ~ "Other",
    TRUE ~ NA_character_   # Handles any other cases or NA values
  ))
  
# Check which rows where altered for testing
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Ethnicity, .after = last_col())

# Sort for testing
  data_with_outliers_removed <- data_with_outliers_removed %>%
  arrange(Ethnicity_reduced)


```

### Create a variable for number of diseases
```{r}
# Create a new column for the number of diseases
# Define the disease columns explicitly
organ_systems <- c("Cancer.Diagnosis.any", "Cardiovascular.any", "Respiratory.Disease.any", 
      "Endocrine.Disease.any", "Skin.Hair.any", "Gastrointestinal.any",
      "Genito.Urinary.any", "Musculoskeletal.any", "Neuropsychological.any", "Reproductive.any","Immune.any")

data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(organ_systems_afflicted_by_disease = rowSums(select(., all_of(organ_systems)), na.rm = TRUE))

# Move all disease columns and the organ_systems_afflicted_by_disease column to the end
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(all_of(organ_systems), .after = last_col()) %>%
  relocate(organ_systems_afflicted_by_disease, .after = last_col())
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(organ_systems_afflicted_by_disease = rowSums(select(., all_of(organ_systems)), na.rm = TRUE))
# Move the organ_systems_afflicted_by_disease column to the end
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(organ_systems_afflicted_by_disease, .after = last_col())

```

### Print colnames: can be removed later
```{r}

# # Using names() with seq_along()
# column_names <- names(data_with_outliers_removed)
# column_indices <- seq_along(data_with_outliers_removed)

# print(column_names)
```

### Harmonize diet
```{r}
# Create harmonized diet variable
# Diet Harmonization
# The following code creates a harmonized diet classification variable by combining information from "Primary.Diet.Past.Year" and "Main.Diet.Factor" fields.
# 
# Harmonized diet categories:
# - "Plant-based": Includes vegan, vegetarian, and Mediterranean diets
# - "Omnivore": Includes mixed/balanced diets, no red meat diets, and pescatarian diets
# - "Carnivore": Exclusively or primarily meat and animal products
# - "Ketogenic": Limited carbohydrate consumption
# - "Other": Cases where diet information is missing or doesn't fit into the above categories
#
# The logic prioritizes "Primary.Diet.Past.Year" information when available, 
# and falls back to "Main.Diet.Factor" when the primary diet information is missing.
data_with_outliers_removed <- data_with_outliers_removed %>%
  mutate(harmonized_diet = case_when(
    !is.na(Primary.Diet.Past.Year) & Primary.Diet.Past.Year %in% c("Vegan. No meat or animal products", 
                                                                 "Vegetarian. No meat products", 
                                                                 "Mediterranean. Emphasis on plant-based products, healthy fats, whole grains, nuts, and seeds. Limited meats, dairy, and processed foods") ~ "Plant-based",
    !is.na(Primary.Diet.Past.Year) & Primary.Diet.Past.Year %in% c("Mixed/Balanced Diet. Includes mixture of fruits, vegetables, meat, fish, dairy, and animal products. No specific restrictions.", 
                                                                "No Red Meat. Typical balanced diet, without any red meat (still consumes fish and poultry)", 
                                                                "Pescatarian. Limits meat products to only fish and seafood") ~ "Omnivore",
    !is.na(Primary.Diet.Past.Year) & Primary.Diet.Past.Year == "Carnivore. Only/majority meat and animal products" ~ "Carnivore",
    !is.na(Primary.Diet.Past.Year) & Primary.Diet.Past.Year == "Ketogenic. Limited carbohydrate consumption" ~ "Ketogenic",
    !is.na(Main.Diet.Factor) & Main.Diet.Factor == "Ketogenic" ~ "Ketogenic",
    !is.na(Main.Diet.Factor) & Main.Diet.Factor %in% c("Meat.Vegetables", "Meat.Fish", "Fish.Vegetables") ~ "Omnivore",
    !is.na(Main.Diet.Factor) & Main.Diet.Factor == "Vegetables" ~ "Plant-based",
    is.na(Primary.Diet.Past.Year) & is.na(Main.Diet.Factor) ~ "Other",
    TRUE ~ "Other"
  ))

# Move the diet columns to the end of the dataframe
data_with_outliers_removed <- data_with_outliers_removed %>%
  relocate(Primary.Diet.Past.Year, Main.Diet.Factor, Main.Diet, harmonized_diet, .after = last_col())

```

### Save Rdata
```{r}
save(data_with_outliers_removed, file = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step3")

```

### Print colnames
```{r}
# Assuming your data frame is named df
colnames(data_with_outliers_removed)
```


# Plot outliers
```{r}
# # Define the output directory
# output_dir <- "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/Visualization of outliers after removal"

# # Create the output directory if it doesn't exist
# if (!dir.exists(output_dir)) {
#   dir.create(output_dir, recursive = TRUE)
# }

# exclude_columns = c("Patient.ID", "PID", "survey_version", "Array", "Collection.Date")

# # Assuming you have a dataframe `data_with_outliers_removed`
# col_numbers <- get_Numeric_and_int_cols(data_with_outliers_removed, exclude_cols = exclude_columns)
# col_names <- names(data_with_outliers_removed)[col_numbers]

# for (col_name in col_names) {
#   # Define the file name for each plot
#   file_name <- paste0(output_dir, "/plot_col_", col_name, ".pdf")
  
#   # Open a PDF device
#   pdf(file_name)
  
#   # Generate the plot
#   plot_with_std_outliers(data_with_outliers_removed, outliers_std_thresh = 5, col_num = which(names(data_with_outliers_removed) == col_name))
  
#   # Close the PDF device
#   dev.off()
# }

```

### Analyze sleep hours distribution across survey versions
```{r}
# # Analyze Hours.of.sleep.per.night distribution by survey version
# cat("=== Distribution of Hours.of.sleep.per.night by Survey Version ===\n")

# for (survey_ver in 1:3) {
#   survey_data <- data_with_outliers_removed %>% filter(survey_version == survey_ver)
  
#   if (nrow(survey_data) > 0) {
#     freq_table <- table(survey_data$Hours.of.sleep.per.night, useNA = "ifany")
#     percentages <- round(prop.table(freq_table) * 100, 2)
    
#     cat("\nSurvey Version", survey_ver, "(n =", nrow(survey_data), "):\n")
    
#     for (i in 1:length(percentages)) {
#       category <- names(percentages)[i]
#       if (is.na(category)) category <- "Missing"
#       cat("  ", category, ": ", percentages[i], "% (n = ", freq_table[i], ")\n", sep = "")
#     }
#   }
# }
```
