---
title: "Imputation"
output: html_document
date: "2024-10-25"
---


#Import functions
```{r}

library(pacman)
p_load(dplyr)


# source("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/set_NA_percent_and_imp_cols().R")
source("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/mixed_correlation_matrix().R")
source("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/imputationFunctions.R")
```

# Data exploration can be removed later
```{r}
# #Load data
# load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# # Rename dataframe in this new script
# data<-data_with_ageDev
# rm(data_with_ageDev) # Remove old dataframe

# print(colnames(data))

```



### Calculate corr matrix
```{r}
  # Remove unnecessary columns
  # data4_without_cols <- subset(data4, select = -c(Hannum.PC, Horvath.PC, Telomere.Values.PC, GrimAge.PC, PhenoAge.PC,
  #                                        SystemsAge.Heart, SystemsAge.Hormone, SystemsAge.Blood, SystemsAge.Brain, 
  #                                        SystemsAge.Inflammation, SystemsAge.Immune, SystemsAge.Kidney, SystemsAge.Liver,
  #                                        SystemsAge.Metabolic, SystemsAge.Lung, SystemsAge.MusculoSkeletal, SystemsAge,
  #                                        OMICmAge, fasting, Main.Diet, NAD, TA65, sulforaphane, DHEA_new, Rapamycin_new,
  #                                        SASP_supressors, Metformin_new, Resveratrol_new, Exosomes, stem_cells, HRT, 
  #                                        spermidine, semaglutide, vitaminD, AKG, Hannum.PCAgeDev, Horvath.PCAgeDev, 
  #                                         GrimAge.PCAgeDev, PhenoAge.PCAgeDev, OMICmAgeAgeDev,
  #                                        SystemsAge.BloodAgeDev, SystemsAge.BrainAgeDev, SystemsAge.InflammationAgeDev,
  #                                        SystemsAge.HeartAgeDev, SystemsAge.HormoneAgeDev, SystemsAge.ImmuneAgeDev,
  #                                        SystemsAge.KidneyAgeDev, SystemsAge.LiverAgeDev, SystemsAge.MetabolicAgeDev,
  #                                        SystemsAge.LungAgeDev, SystemsAge.MusculoSkeletalAgeDev, SystemsAgeAgeDev, DunedinPACE))

# OMICage och grimage är ej korrolerade över 0,8
# 
# Kolla OMICage-dunedin och dunedin-grimAge

# corr_df<-mixed_correlation_matrix(data4, cols_to_skip = c("Patient.ID", "PID", "Collection.Date", "Array", "survey_version"))

```


### Impute survey 1 - All participants
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# analyze_column(data, "Menopause")

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")

impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1")

```

# Do imputation for gestational diabetes 
```{r}

print("Running gestation diabetes")
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

data <- data %>% filter(Gestational.Diabetes == 1)


# Separate columns to keep unimputed (because will not be used in the later analysis) (due to no missing data)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")

impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_gestational_diabetes")


```



# Do imputation for diabetics
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

data <- data %>% filter(Diabetes2 == 1)

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")


impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_diabetes")


```


# Do imputation for prediabetics 
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

data <- data %>% filter(Prediabetes == 1)
# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")


impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_prediabetics")


```



# Do imputation for non-diabetics non prediabetics 
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

data <- data %>% filter(Prediabetes == 0 & Diabetes2 == 0)
# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")


impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_healthy")


```

# Do age-wise imputation
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# Calculate the median age
median_age <- median(data$Decimal.Chronological.Age, na.rm = TRUE)
print(paste("median:", median_age))

# Split the data into two data frames
data_lower <- data[data$Decimal.Chronological.Age < median_age, ]
data_upper <- data[data$Decimal.Chronological.Age >= median_age, ]

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")

#Run imputation for older group
impute_survey_data(data_to_drop_cols_from=data_upper, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_older")

#Run imputation for younger group
impute_survey_data(data_to_drop_cols_from=data_lower, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_younger")

```



### Impute survey 1 - only females
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# Remove males from data
data <- data %>% filter(Biological.Sex == "Female")
# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")


impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_females")


```

### Impute survey 1 - only males
```{r}

#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# Remove females from data
data <- data %>% filter(Biological.Sex == "Male")

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Hours.of.sleep.per.night",
  "Tobacco.Use", "Exercise.per.week", "Drug.Alcohol.mother", "Drug.Alcohol.father", "Mother.Nicotine.Use", "Mother.Pregnancy.Complications", "Caffeine.Use", "Marital.Status", "Sexual.Frequency")

impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 1, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_males")


```

### Impute survey 2 - All participants
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Tobacco.Use")

impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 2, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey2")

```


### Impute survey 3 - All participants
```{r}
#Load data
load("C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/data_after_step4")
# Rename dataframe in this new script
data<-data_with_ageDev
rm(data_with_ageDev) # Remove old dataframe

# Separate columns to keep unimputed (because will not be used in the later analysis)
cols_to_exclude <- c("Alcohol.per.week", "Level.of.Education", "Tobacco.Use", "Hours.of.sleep.per.night")

impute_survey_data(data_to_drop_cols_from=data, surveyVersion = 3, missing_threshold = 50, 
                                 amount_of_mice_datasets_to_impute = 2, max_iterations_per_dataset = 1, 
                                 cols_to_exclude_from_imputation_entirely = cols_to_exclude, savePath = "C:/Users/danie/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey3")

```

