```{r}
# import functions: note the file has been renamed to remove parentheses
source("C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Program/run_models_survey1.R")
```

# Generate main models
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor", "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age", "Biological.Sex"),
                     suffix = "_main", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1")
```

Generate sensitivity analysis for diabetics
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age", "Biological.Sex"), suffix = "_only_diabetics", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_diabetes2") 
```

Generate sensitivity analysis for prediabetics
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age", "Biological.Sex"), suffix = "_only_prediabetes", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_prediabetes") 
```

Generate sensitivity analysis for non diabetic type 2 and non prediabetic
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age", "Biological.Sex"), suffix = "_only_healthy", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_healthy") 
```

Generate sensitivity analysis for males
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age"), suffix = "_only_males", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_males") 
```

Generate sensitivity analysis for females
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age"), suffix = "_only_females", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_only_females") 
```

Generate sensitivity analysis for old
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age"), suffix = "_only_old", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_older") 
```

Generate sensitivity analysis for young
```{r}
#Generate Sensitivity analysis for excluding exercise
run_models_survey1(Anti_aging_interventions = c("Rapamycin_new", "Metformin_new", "fasting", "NAD", "TA65", 
                     "sulforaphane", "DHEA_new", "SASP_supressors", "Resveratrol_new", "Exosomes", 
                     "HRT", "spermidine"), 
                     covariates = c("Alcohol_per_week_numeric",  
                                    "Education_levels_numeric", "Stress.Level", "Tobacco.Use.Numeric", 
                                    "Main.Diet.Factor" , "BMI", "Caffeine.Use_numeric", 
                                    "Marital.Status_numeric", "Sexual.Frequency_numeric", "Hours.of.sleep.per.night_numeric"), 
                     cov_to_always_include = c("Decimal.Chronological.Age"), suffix = "_only_young", 
                     data_path = "C:/Users/danwik/OneDrive - Karolinska Institutet/Documents/Project 2 - Vscode/Output/imputed_survey1_younger") 
```


